---
title: "Voting Frequencies and Democratic Equity: Exploratory Data Analysis "
author: "Elizabeth Deng, Judy Lee, Naiyyra Abdel Aziz, Sukhman Dhillon"
date: "11/14/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
options(repos = list(CRAN="http://cran.rstudio.com/"))
library("dplyr")
library("knitr")
library("ggplot2")
```

```{r echo=FALSE}
source("../source/summary_info.R")
```

## Summary 
The average percentage of white voters in the 2018 election across the United States was `r summary_info$avg_white_pct`%.
The average percentage of nonwhite voters in the 2018 election across the United States was `r summary_info$avg_nonwhite_pct`%.
The state with the highest average median household income of voters in the 2018 election was `r summary_info$max_hh_inc`.
The average percentage of female voters in the 2018 election across the United States was `r summary_info$avg_female_pct`%.
The state with the highest average percentage of voters that were aged 29 and under in the 2018 election was `r summary_info$max_young`.

## Table  
### An Overview of Demographics (2016 Presidential Election)  
The table below shows aggregate information about the states and how their populations  
voted in the 2016 Presidential Election. All numerical value refer to percentages.   
For Columns 4-10, these refer to each demographic as a percentage of *total* population.  

## Why This Is Relevant  
We chose to include this table because it gives a brief glimpse into how different   
demographics within a state can affect the candidate it votes for. For example,   
states with high (over 60%) rural populations tended to vote overwhelmingly for   
Trump; meanwhile, states with higher median household incomes chose Clinton. 

```{r pres_16_table, echo=FALSE}
source("../source/2018_analysis.R")
kable(pres_16_table)
```

 

```{r echo=FALSE}
source("../source/p2_map.R")
```

## Map
```{r echo=FALSE}
complete_map
```

## Purpose
We included this chart because we noticed that in the dataset, the type of ballot
box was split into two categories: drop boxes and voting centers. Because of this,
we thought it would be interesting to see what the distribution of drop boxes and
voting centers across Washington state were like to see if any obvious patterns arise,
like whether location affects the type of ballot box is near someone.

## Observations
After creating this map, something that can be seen quickly is the high number of
data points near Seattle. Specifically, these data points are blue which represent 
drop boxes. We can learn from this chart that more drop boxes are put in more densely
populated areas. However, that means that areas with less people have less options
for where they can return their ballots. This can make it harder for these populations
to vote and unfairly affect voting rates. Something else we can learn from this map
is that there's almost one voting center in each county. However, some counties are
larger than others which means that if someone in a larger county is interested in
voting in person, they have to go a further distance than a smaller county. This can
also limit voting rates in these populations.

## Scatterplot
``` {r echo=FALSE}
source("../source/trump_percent_scatterplot.R")
```

```{r echo=FALSE}
scatter1
```

## Purpose

We included this chart because as per our research question, we want to analyze
how household income correlates with the type of candidate they vote for. We 
want to observe if there are, or aren't, any trends along all the counties in 
America, which is what made this dataset perfect as it has the statistics and data
of each county and state, and the number of votes in each county for each candidate. 
This information can not only help with predicting voting trends, but understanding 
the finanical setting of individuals and how it infleunces their voting decisions.


## Observations
This graph revealed a trend between house hold income and the percentage of people
that would vote for Trump in the Presidential Election. In almost all counties in America.The largest cluster of votes for Trump were individuals with a median household income of between 40,000 - 50,000 dollars a year. This reveals that people in lower social economic statuses tended to vote for Trump. 

## Bar chart

```{r echo=FALSE}
source("../source/Nov2020_Bar.R")
```

```{r echo=FALSE}
MF_bar
```

## Purpose

The purpose of this chart is to compare the percentage of votes between gender in the United States in November 2020. The distribution of this data set is into many different categories, such as gender, race, Hispanic origin, total percentage of registration, total percentage of votes, and more. For that reason, we thought it would be interesting to make a distribution in a bar chart that shows the comparison of votes in gender by the state in the United States. 


## Observations

Through this chart, we can determine the gender participation in voting by states in the United States. For example, we can notice that the District of Columbia had the highest voting participation, and the females voted more, with 79.5 percent, while 75.5 percent of males voted. Also, we can observe that Arkansas had the lowest participation, and the females voted more, with 54.1 percent and 54.1 percent of males voting. Overall, females have participated more than males in voting. 
